---
title: "DataAnalyses_v2"
author: "sbsambado"
date: "6/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(car)
library(knitr)
library(lme4)
library(jtools)
library(effects)
library(psych)
library(gt)
library(MuMIn)
library(kableExtra)

Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")
dim(Tick) # 1449   11


```

### Tick abundances

```{r}

sum(Tick$total, na.rm = TRUE) # 2692 total ticks

## species collection

## ticks by year

Tick %>%
  group_by(year) %>%
  dplyr::select(year, total) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd))


## tick by pre- and post-fence

# pre
pre <- Tick %>%
  filter(year == 2016) %>%
  group_by(site) %>%
  dplyr::select(site, total) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd))

# post

post <- Tick %>%
  filter(year != 2016) %>%
  group_by(site) %>%
  dplyr::select(site, total) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd))

# wilcoxon rank sum test
wilcox.test(pre$total_mean, post$total_mean)

## ticks by plot

Tick %>%
  group_by(plot) %>%
  dplyr::select(plot, total) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd))


## ticks by site

Tick %>%
  group_by(site) %>%
  dplyr::select(site, total) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd))


## ticks by month

# deoc
Tick %>%
  group_by(month) %>%
  drop_na() %>%
  summarise_each(funs(mean, sd)) %>%
  dplyr::select(month, deoc_mean, deoc_sd)


# ixodes

  Tick %>%
  group_by(month) %>% # 
  drop_na() %>%
  summarise_each(funs(mean, sd)) %>%
  dplyr::select(month, ipac_mean, ipac_sd)
```

### D. occidentalis counts response to cattle (plot) and climate (site)

To test the impact of post fence, I only run regression on D. occidentalis for 2017, 2018, and 2019.
```{r}
## get data right

tick_no2016 <- Tick %>%
  drop_na() %>%
  filter(year != "2016")

# re-order levels to make it easier to interpret effect of fence

tick_no2016$plot <- factor(tick_no2016$plot,
                             levels = c("Partial","Open",  "Total"))


tick_no2016$site <- factor(tick_no2016$site,
                             levels = c("Intermediate", "Arid", "Mesic"))


## best fit model
mod_deoc_no2016 <- glmmTMB(deoc ~ plot + site + plot*year + (1|month),
                           family = nbinom2,
                           data = tick_no2016)

## model diagnostics

# summary
summary(mod_deoc_no2016)

# R2
r.squaredGLMM(mod_deoc_no2016)
  # marginal = 0.11
  # conditional = 0.88


#model diagnostics
Validation <- function(bestmodel, tick_no2016){
  Model.Residuals <- residuals(bestmodel, type = 'pearson')
  Model.Fitted <- fitted(bestmodel)
  scatter.smooth(Model.Fitted ~ Total)
  scatter.smooth(Model.Residuals~Model.Fitted)
  hist(Model.Residuals)
  qqnorm(Model.Residuals)
  Validate <- cbind(tickdata, Model.Residuals)
  boxplot(Model.Residuals~site)
  boxplot(Model.Residuals~plot)
  boxplot(Model.Residuals~(plot*year))
  boxplot(Model.Residuals~month)
}



## predict plot
# Estimated abundance ignoring correlation. Points represent plot average counts. Error bars represent the 95% Wald-type confidence intervals for the predicted average count

pred <- glmmTMB(deoc ~ plot + site, data = tick_no2016, family = nbinom2)
newdata0 = newdata = unique(tick_no2016[,c("plot","site")])
temp = predict(pred, newdata, se.fit = TRUE, zitype = "response")
newdata$predFE = temp$fit
newdata$predFE.min = temp$fit-1.98*temp$se.fit
newdata$predFE.max = temp$fit+1.98*temp$se.fit

real = ddply(tick_no2016, ~ plot + site, summarize, m = mean(deoc))
newdata$plot <- factor(newdata$plot,
                       levels = c("Total", "Partial", "Open"),
                       labels = c("- Wildlife & Cattle", "Control", "+ Wildlife & Cattle"))
ggplot(newdata, aes(plot, predFE, color = site)) +
  geom_point(size = 3.5) +
  geom_errorbar(aes(ymin = predFE.min, ymax=predFE.max), width = .1, size = 1.5) +
  #geom_point(data=real, aes(x=plot, y=m)) +
  labs(y = "Estimated Average Tick Abundance", x = "", color = "Site") +
  theme_classic() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(face = "bold", size = 14, color = "black"),
        legend.position = c(.83,.8),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold"),
        legend.background = element_rect(color = "black", fill = "white", size = .1),
        legend.title.align = .5) + #,
    scale_color_manual(values =c('tan1','brown1','blue'))

```

## Supplement

### D. occidentalis

Variable selection
```{r}
mod_0_plotID <- glmmTMB(deoc ~ (1|plotID),
                 data = tick_no2016,
                 family = "poisson")

mod_0_month <- glmmTMB(deoc ~ (1|month),
                 data = tick_no2016,
                 family = "poisson")

mod_0_plotIDmonth <- glmmTMB(deoc ~ (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "poisson")

AICctab(mod_0_plotID, mod_0_month, mod_0_plotIDmonth) # mod_0_plotIDmonth best model



mod_1 <- glmmTMB(deoc ~ site + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")

mod_2 <- glmmTMB(deoc ~ plot + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")

mod_3 <- glmmTMB(deoc ~ year + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")

mod_4 <- glmmTMB(deoc ~ site + plot + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")

mod_5 <- glmmTMB(deoc ~ site + plot + year + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")


mod_6 <- glmmTMB(deoc ~ site + plot + plot*year + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")

mod_7 <- glmmTMB(deoc ~ site + plot + plot*site + (1|plotID) + (1|month),
                 data = tick_no2016,
                 family = "nbinom2")


AICctab(mod_1, mod_2, mod_3, mod_4, mod_5, mod_6, mod_7) 
# w/o random effects, mod_7 best
# w/ random effects, mod_6 is best ... but single random effect month performs better than them together


# plot_summs(mod_1, mod_2, mod_3, mod_4, mod_5,  mod_7) # removed mod_6 to make more interprettable 



library(lmtest)
# lrtest(nested, complex) # p > 0.05 so don't use complex model
```


Explore best fit model type
```{r}

mod_poisson <-  glmmTMB(deoc ~ plot + site + plot*year + (1|month),
                           family = "poisson",
                           data = tick_no2016)

mod_nb1 <-  glmmTMB(deoc ~ plot + site + plot*year + (1|month),
                           family = nbinom1,
                           data = tick_no2016)

mod_nb2 <-  glmmTMB(deoc ~ plot + site + plot*year + (1|month),
                           family = nbinom2,
                           data = tick_no2016)


AICtab(mod_poisson, mod_nb1, mod_nb2) # mod_nb2 clearly best fit model w/ dAIC 0 compared to mod_nb1 dAIC 46
```

## All ticks

Set up
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(car)
library(knitr)
library(lme4)
library(jtools)
library(effects)
library(psych)
library(gt)
library(MuMIn)
library(kableExtra)


```

*Notes*
plot = cattle
site = climate

## Question 1: How do environmental stressors - cattle & climate - impact tick counts? 

--> best fit model incorporates both stressors
--> however, looked at individually, climate is the more strong predictor of tick counts
--> best fit random effects were month & plotID  

Step 1. Import data set
```{r import}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")
```

Step 2. Build best fit models
- look at Tejon_Analyses_20211021 for all model combinations

```{r best fit models}
mod_plot <- glmer.nb(total ~ plot + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")

mod_site <- glmer.nb(total ~ site  + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")

mod_both <- glmer.nb(total ~ site + plot + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")


model_results <- AIC(mod_plot, mod_site, mod_both)

model_models <- list(mod_plot, mod_site, mod_both)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)
for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$variables <- c("plot", "site", "plot + site")

model_models <- kable(model_results, digits = 2, align = "c") %>%
  kable_styling()

model_models

```

Step 3. Model diagnostics

3A. summary functions
```{r model diagnostics pt 1}

## summary()
summary(mod_plot) # p = 0.00770
summary(mod_site) # p = 0.00567
summary(mod_both) # p = 0.000187

## summ() function
               # Est    S.E.    z value   p
summ(mod_plot) # -2.05   0.77    -2.67   0.01  ; Pseudo-R² (total) = 0.81 
summ(mod_site) # -2.09   0.75    -2.77   0.01  ; Pseudo-R² (total) = 0.81 
summ(mod_both) # -2.82   0.76    -3.74   0.00  ; Pseudo-R² (total) = 0.81

```

3B. Akaike weight
- get relative support for each model by standardizing the likelihood of the model
```{r model diagnostics pt 2}

## Option 1: Calculate weights manually

# calculate deltai2

delta1 <- exp(-0.5*(AIC(mod_plot) - AIC(mod_both)))
delta2 <- exp(-0.5*(AIC(mod_site) - AIC(mod_both)))
delta3 <- exp(-0.5*(AIC(mod_both) - AIC(mod_both)))

# sum likelihoods
sum_likes <- sum(c(delta1, delta2, delta3))

# calculate weights
weight1 <- delta1/sum_likes
weight2 <- delta2/sum_likes
weight3 <- delta3/sum_likes

# lets make a table
weights <- tribble(
  ~Model,   ~Delta,           ~Weights,   ~Variable,
  "mod_plot", round(delta1, 2), round(weight1, 2), "plot",
  "mod_site", round(delta2, 2), round(weight2, 2), "site",
  "mod_both", round(delta3, 2), round(weight3, 2), "both") %>% 
  gt() %>% 
  tab_style(
    style = cell_text(size = "small"),
    locations = cells_body())


weights


## Option 2: dredge the model

min(Tick$total)
model <- na.omit(Tick)
big.model <- glm(total  ~ plot + site + plot*site, 
                 family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  tab_header(title="Model Selection",
             subtitle = "total ~ plot + site + plot*site")

glm.best

# delta = 0 is best model

```

Step 4. Plot model outputs
```{r plot model outputs, results = 'asis'}

## plot_summs()
plot_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           model.names = c("Only Cattle (plot)", "Only Climate (site)", "Both Cattle & Climate"),
           inner_ci_level = .8,
           robust = TRUE,
           coefs = c("Partial plot" = "plotPartial",
                     "Total plot" = "plotTotal",
                     "Inter. site" = "siteIntermediate", 
                     "Mesic site" = "siteMesic"),
           legend.direction = "horizontal")

```

Step 5. Create output table of model results
```{r model table statistics}
export_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           error_format = "({statistic}, 95% CI [{conf.low}, {conf.high}], p = {p.value})",
             to.file = "docx", 
             file.name = "Tejon_BestFitModels.docx")
```


## Question 2: How do environmental stressors impact populations of two important vertebrate hosts -- lizards and mammals -- in the tick system?

--> lizards are significantly more impacted by climate than plot
--> mammals are trending to be significantly impacted by plot than climate

Step 1. Import & tidy dataset
```{r import and tidy}

pop <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mamliz_popestimates.csv")

# rename columns to be consistant
names(pop)[5] <- "mam_estimate"

# change data structure for analyses
pop$site <- as.factor(pop$site)
pop$plot <- as.factor(pop$plot)

```

Step 2. Compare means of population estimates

2A. Lizards
```{r lizards}
# lizards were only sampled at two sites, so will use a two sample t-test if normal

# check normality
qqPlot(pop$liz_estimate) # looks normal, but..
shapiro.test(pop$liz_estimate) # p < 0.05, so data are not normal

# check variance
leveneTest(pop$liz_estimate ~ pop$site) #  p < 0.05, F = 7.4; variances are different

# non-parametric Mann-Whitney U is testing the shape of the two distributions are different

# wilcox rank sum test with continuity correction (non-equal variances)
wilcox.test(liz_estimate ~ site, data = pop) #W = 11, p-value  = 0.01306

# p < 0.05, reject H0 that means are equal to 0

## lizard population estimates are different at Intermediate and Arid sites

```

2B. Mammals
```{r mammals}

# mammal populations did not significantly differ across sites or plots but that may be due to low sample numbers with kruskal wallis

# let's do open vs total plots --> mann-whitney U test

pop_sub <- subset(pop, pop$plot != "Partial")
wilcox.test(mam_estimate ~ plot, data = pop_sub)
# W = 11, p-value = 0.05488 (trending significant)
```

## Question 3: Are tick counts impacted by lizard and mammal population estimates?

--> lizards population estimates are more associated with tick counts than mammals

Step 1. Import & tidy dataset
```{r import and tidy}
FINAL <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MixedModels_Dataset.csv")

# drop N/A values
FINAL <- FINAL[which(FINAL$mam_estimates != "#N/A"),]
FINAL <- FINAL[which(FINAL$liz_estimates != "#N/A"),]

# check it worked
unique(FINAL$mam_estimates)
unique(FINAL$liz_estimates)

str(FINAL)
```

Step 2. Spearman's rank correlation

```{r spearman rank correlation}
# select subset of data to use for test & visualize
FINAL_cor <- subset(FINAL, select = c(6,20:21))

# change structure of data for analyses
FINAL_cor$total <- as.numeric(FINAL_cor$total)
FINAL_cor$mam_estimates <- as.numeric(FINAL_cor$mam_estimates)
FINAL_cor$liz_estimates <- as.numeric(FINAL_cor$liz_estimates)

# visualize data
pairs.panels(FINAL_cor)

## run correlation test

# lizards
cor.test(FINAL_cor$total,FINAL_cor$liz_estimates,
         method = "spearman")
# p-value = 1.755e-05
# rho = 0.1687151

# mammals
cor.test(FINAL_cor$total,FINAL_cor$mam_estimates,
         method = "spearman")
# p-value = 0.3669
# rho = -0.03569831 


```

Step 3. Build models : GLM with Poisson distribution 

~ I think this may not be appropriate based on the aggregate of the pop estimates ~

```{r glm with poisson}

mod_mam <- glm(total ~ mam_estimates, data = FINAL,
               family = "poisson")

mod_liz <- glm(total ~ liz_estimates, data = FINAL,
               family = "poisson")

mod_ml <- glm(total ~ mam_estimates + liz_estimates, data = FINAL,
               family = "poisson")

## Model comparison

model_results <- AIC(mod_mam, mod_liz, mod_ml)

model_models <- list(mod_mam, mod_liz, mod_ml)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)

for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$variables <- c("mam", "liz", "mam + liz")

model_models <- kable(model_results, digits = 2, align = "c") %>%
  kable_styling()

model_models
```

Step 4. Model diagnostics

4A. summary functions
```{r model diagnostics pt 1}

## summary()
summary(mod_mam) # p = <2e-16
summary(mod_liz) # p = <2e-16
summary(mod_ml) # p = <2e-16

## summ() function
               # Est    S.E.    z value   p
summ(mod_mam) #  0.91   0.04    23.12   0.00   
summ(mod_liz) # -1.46   0.09   -17.04   0.00 
summ(mod_ml) #  -1.26   0.10   -12.49   0.00

```

4B. Dredge model
```{r model diagnostics pt 2}
## Option 2: dredge the model

min(FINAL$total)
model <- na.omit(FINAL)
big.model <- glm(total ~ mam_estimates + liz_estimates, 
               family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  tab_header(title="Model Selection",
             subtitle = "total ~ mam + liz")

glm.best

```



