---
title: "Data Figures and Tables"
author: "sbsambado"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(ggfortify)
library(gtsummary)
library(lme4)
library(effects)
library(jtools)
library(huxtable)
library(MuMIn)
library(gt)
library(plyr)
library(MASS)
library(glmmTMB)

# set up some visualization vectors
clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('brown1','tan1','lightgoldenrod1')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))

ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))

xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))


```

## Figure 1
**Tick counts by one variable (plot or site)**

```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

# omit 0s to make plots look nicer, but this may not be kosher
Tick_no0 <- Tick[which(Tick$total != 0),]

plot1 <- ggplot(Tick_no0, aes(x = plot, y = log(total +1), fill = plot))+
  #geom_jitter(alpha = .15, width = .2)+
  geom_boxplot()+
  theme_bw() +
  labs(x = 'Plot', y = 'Log of tick counts (2016 - 2019)') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  guides(fill = FALSE)+
  annotate('text', label='A', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7) +
  scale_x_discrete(breaks = c("Open", "Partial", "Total"),
                   labels = c("C", "+WC", "-WC"))

site1 <-ggplot(Tick_no0, aes(x = site, y = log(total +1), fill = site))+
  #geom_jitter(alpha = .15, width = .2)+
  geom_boxplot() +
  theme_bw() +
  labs(x = 'Site', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)+
  annotate('text', label='B', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

var_plot <-ggarrange(plot1, site1 + rremove('ylab'), # + rremove('ylab')
          ncol = 2)
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig1_OneVarTick_v2.jpeg", dpi = 320)    

## changed this about figure to make sure y labels were consistent
annotate_figure(var_plot,
          left = text_grob("Log of tick counts (2016 - 2019)", rot = 90, face = "bold")) 

      


```

## Figure 2. 
** Tick counts by two variables (plot and site)**

```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

# omit 0s to make plots look nicer, but this may not be kosher
Tick_no0 <- Tick[which(Tick$total != 0),]


Tick_no0$plot <- recode(Tick_no0$plot,
                        Open = "+WC",
                        Partial = "C",
                        Total = "-WC")
ggplot(Tick_no0, aes(x = site, y = log(total + 1), fill = plot))+
  geom_boxplot()+
  #geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Site', y = "Log of tick counts (2016 - 2019)", fill = 'Plot') +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = c(.911, .816), # horizontal, verticle
        legend.box.background = element_rect(color = 'black', fill =NA),legend.key.size = unit(.8,"line")) +  
  annotate('text', label='C', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig2_TwoVarTick_v2.jpeg", dpi = 320)  

```

## Figure 3. 
**Climate by site**

```{r}
FINAL <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MixedModels_Dataset.csv")

# change data structure
FINAL$site <- as.factor(FINAL$site)
FINAL$plot <- as.factor(FINAL$plot)
FINAL$tempdifC <- as.numeric(FINAL$tempdifC)
FINAL$mean_maxC <- as.numeric(FINAL$mean_maxC)

ggplot(FINAL, aes(y = tempdifC, x = site, fill = site))+
  geom_boxplot() +
  #geom_jitter(alpha = .15, width = .2)+
  theme_bw()+
  labs(x = 'Site', y = 'Difference in monthly\n mean temperature (°C)', fill = "Site")+
  xtitletheme + ytitletheme +
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)
#ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig3_DiffTempCSite.jpeg", dpi = 320)

```
## Figure 4. 
**Effects plot: tick counts ~ cattle + climate**

```{r}
## get data right

tick_no2016 <- Tick %>%
  drop_na() %>%
  filter(year != "2016")

# re-order levels to make it easier to interpret effect of fence

tick_no2016$plot <- factor(tick_no2016$plot,
                             levels = c("Partial","Open",  "Total"))


tick_no2016$site <- factor(tick_no2016$site,
                             levels = c("Intermediate", "Arid", "Mesic"))


## best fit model
mod_deoc_no2016 <- glm.nb(deoc ~ plot + site + plot*year,
                       #family = "negbinomial",
                           data = tick_no2016)


## plot_summs()
exp_estimates <- plot_summs(mod_deoc_no2016,
           scale = TRUE,
           #model.names = c("Only Cattle (plot)", "Only Climate (site)", "Both Cattle & Climate"),
           inner_ci_level = .8,
           #center = TRUE,
           coefs = c("Year"= "year",
                    
                     "Partial plot" = "plotPartial",
                     
                     
                      "Inter. site" = "siteIntermediate",
                     
                     
                    
                     
                     
                    
           "-WC" = "plotTotal",
            "+WC * Year"="plotOpen:year",
            "-WC * Year"="plotTotal:year",
           "+WC"="plotOpen",
            "Mesic" = "siteMesic",
           "Arid" = "siteArid"),
           exp = TRUE,
           colors = "CUD")

apatheme = theme_classic() +
  theme(axis.text = element_text(face = "bold", size = 12, color = "black"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(face = "bold", size = 12, color = "black"))
exp_estimates + apatheme
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig4_Coefficients_v2.jpeg", dpi = 320)




```
### Supplemental

## Figure 5.
**Tick counts by month (site)**

```{r}

Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Tick$site <- factor(Tick$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

Tick$plot <- factor(Tick$plot, 
                    levels = c('Open', 'Partial', 'Total'))

ggplot(Tick, aes(x = site, y = log(total +1), fill = site))+
  facet_wrap(~ month, ncol = 6)+
  geom_boxplot()+
  theme_bw() +
  scale_x_discrete(labels = c("A", "I", "M")) +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'))+
  labs(x = 'Site', y = 'Log of tick counts (2016 - 2019)') +
  xtitletheme + ytitletheme +
  guides(fill = FALSE)+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))
#ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig5_MonthlyTickSite.jpeg", dpi = 320)



```

## Figure 6. 
**Tick counts by month (both variables)**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Tick$site <- factor(Tick$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))


Tick$plot <- recode(Tick$plot,
                        Open = "+WC",
                        Partial = "C",
                        Total = "-WC")


ggplot(Tick, aes(x = site, y = log(total +1), fill = plot))+
  facet_wrap(~ month, ncol = 6)+
  geom_boxplot()+
  theme_bw() +
  scale_x_discrete(labels = c("A", "I", "M")) +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'),
        legend.position =  'bottom')+
  labs(x = 'Site', y = 'Log of tick counts (2016 - 2019)', fill = "Plot") +
  xtitletheme + ytitletheme +
  #guides(fill = FALSE)+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig6_MonthlyTickSitePlot_v2.jpeg", dpi = 320)
```

## Figure 7.
**Vertebrate population estimates**

```{r}
pop <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mamliz_popestimates.csv")

# rename columns to be consistant
names(pop)[5] <- "mam_estimate"

# change data structure for analyses
pop$site <- as.factor(pop$site)
pop$plot <- as.factor(pop$plot)

liz <- ggplot(pop, aes(x = site, y = liz_estimate, fill = site))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  labs(x = "Site", y = "Lizard Pop. Estimates")+
  guides(fill = FALSE)+
  annotate('text', label='A', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)


# omit outlier
pop_sub <- pop[which(pop$mam_estimate != "216"),]

pop_sub$plot <- recode(pop_sub$plot,
                        Open = "+WC",
                        Partial = "C",
                        Total = "-WC")

mam <- ggplot(pop_sub, aes(x = plot, y = mam_estimate, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  labs(x = "Plot", y = "Mammal Pop. Estimates")+
  guides(fill = FALSE)+
  annotate('text', label='B', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)


ggarrange(liz, mam)
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig7_VertPopEstimates_v2.jpeg", dpi = 320)

```
## Figure 8. 
**Dredge of best fit tick count models**
```{r}

Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

min(Tick$total)
model <- na.omit(Tick)
big.model <- glm(deoc  ~ plot + site + plot*year, 
                 family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  fmt_number(columns = c(1,4,7:10),
             decimals = 2) %>%
  tab_header(title="Model Selection",
             subtitle = "deoc count ~ plot + site + plot*year") 

#glm.best %>%
#  gtsave("data_visualizations/Fig8_ModelSelection.png")
```

## Figure 9
**Model outputs for GLMM negative binomial**
```{r}
# export_summs(mod_plot, mod_site, mod_both,
#            scale = TRUE,
#            error_format = "{statistic}, 95% CI [{conf.low}, {conf.high}]",
#            borders = .5,
#   outer_borders = 0,
#              to.file = "pdf", 
#              file.name = "data_visualizations/Fig9_BestFitModelsQ1.pdf")

```

## Figure 10.
**summaries of tick counts by site and plot**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$plot <- recode(Tick$plot,
                        Open = "+WC",
                        Partial = "C",
                        Total = "-WC")

counts <- Tick %>%
  group_by(site, plot) %>%
  tally(total) 


pdf("data_visualizations/Fig10_TickCount_v2.pdf")
grid.table(counts)

```

## Figure 11.
**summaries of tick counts by month**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")
Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))

counts_month <- Tick %>%
  group_by(month) %>%
  tally(total) 


#pdf("data_visualizations/Fig11_TickCountMonth.pdf")
#grid.table(counts_month)
```

## Figure 12.
**Counts by tick species and month (Table)**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))


Tick_wide <- gather(Tick, species, count, deoc:other, factor_key = TRUE)

Tick_wide$species <- gsub("deoc", "Dermacentor occidentalis",Tick_wide$species )

Tick_wide$species <- gsub("deva", "Dermacentor variabilis",Tick_wide$species )

Tick_wide$species <- gsub("ipac", "Ixodes pacificus",Tick_wide$species)

unique(Tick_wide$species)

Tick_wide$count <- as.numeric(Tick_wide$count)

ipac_count <-Tick_wide %>%
  group_by(month,species) %>%
  filter(species == "Ixodes pacificus") %>%
  tally(count) %>%
  dplyr::select(month, n)

deoc_count <-Tick_wide %>%
  group_by(month,species) %>%
  filter(species == "Dermacentor occidentalis") %>%
  tally(count) %>%
  dplyr::select(month, n)

deva_count <-Tick_wide %>%
  group_by(month,species) %>%
  filter(species == "Dermacentor variabilis") %>%
  tally(count) %>%
  dplyr::select(month, n)

derm_ticks <- join(deoc_count, deva_count, by = "month")

full_ticks <- join(ipac_count, derm_ticks, by = "month")

names(full_ticks) <- c("Month", "IPAC", "DEOC", "DEVA")



#pdf("~/Google Drive/Tejon_clean/data_visualizations/Fig12_TickCountMonthSpecies.pdf")
#grid.table(full_ticks)
```


## Figure 13.
**Tick counts by months (Graph)**

```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))


Tick_wide <- gather(Tick, species, count, deoc:other, factor_key = TRUE)

Tick_wide$count <- as.numeric(Tick_wide$count)

species_count <-Tick_wide %>%
  group_by(month,species) %>%
  filter(species != "other") %>%
  tally(count) 

species_count$species <- factor(species_count$species, 
                                levels = c("ipac", "deoc", "deva"), 
                                labels = c("I. pacificus", "D. occidentalis", "D. variabilis"))

ggplot(species_count, aes(month, n)) +
  geom_histogram(stat = "identity") +
  facet_wrap(~species, scales = "free_y", ncol = 1)  +
  labs(x = "Month", y = "Count") +
  scale_x_discrete(labels = c("J", "F", "M","A",
                              "M","J","J","A",
                              "S", "O", "N", "D"))+
  theme_bw() +
  theme(strip.background = element_rect(fill = "grey95"),
        strip.text = element_text(size=7, face = 'bold'),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7)) +
   ytitletheme + xtitletheme 
#ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig13_TickCountsMonthSpeciesGraph.jpeg", dpi = 320)

  


```

## Figure 14.
**Raw mammal counts**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")


Tick$month <- factor(Tick$month, 
                         levels = c('January', 'February', 'March',
                                    'April', 'May', 'June', 'July',
                                    'August', 'September', 'October',
                                    'November', 'December'))

 Tick_wide <- gather(Tick, species, count, deoc:other, factor_key = TRUE)
 
Tick_wide$count <- as.numeric(Tick_wide$count)

species_count <-Tick_wide %>%
   group_by(month,species) %>%
   filter(species != "other") %>%
   tally(count) 
 
species_count$species <- factor(species_count$species, 
                                 levels = c("ipac", "deoc", "deva"), 
                                 labels = c("I. pacificus", "D. occidentalis", "D. variabilis"))
 
ggplot(species_count, aes(month, n)) +
   geom_histogram(stat = "identity") +
   facet_wrap(~species, scales = "free_y", ncol = 1)  +
   labs(x = "Month", y = "Count") +
   scale_x_discrete(labels = c("J", "F", "M","A",
                               "M","J","J","A",
                               "S", "O", "N", "D"))+
   theme(strip.background = element_rect(fill = "gray95"),
         strip.text = element_text(size=6, face = 'bold')) +
   theme(axis.text.x = element_text(size = 4),
         axis.text.y = element_text(size = 4))+
   theme_bw() +  ytitletheme + xtitletheme
#ggsave("data_visualizations/Fig14_TickCountsMonthSpeciesGraph.jpeg", dpi = 320)
```

## Figure 15.
**Tick counts by year and plot**
```{r}
Tick <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$plot <- recode(Tick$plot,
                        Open = "+WC",
                        Partial = "C",
                        Total = "-WC")

Tick_aggregate <- Tick %>%
   group_by(total, year, plot, site) %>%
   tally(total)


ggplot(Tick_aggregate, aes(plot, n,fill = site)) +
   geom_histogram(stat = "identity")+
   facet_wrap(~year, nrow = 1)  +
   labs(x = "Plot", y = "Tick count", fill = "Site") +
   scale_fill_manual(values = c('brown1','tan1','lightgoldenrod1'))+
   theme_bw() +
   theme(strip.background = element_rect(fill = "grey95"),         strip.text = element_text(size=7, face = 'bold'),
         axis.text.x = element_text(size = 6),
         legend.position = "bottom") +
    ytitletheme + xtitletheme 
ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig15_TickCountsYearPlot_v2.jpeg", dpi = 320)
```

## Figure 16
**Estimated abundance ignoring correlation. Points represent plot average counts. Error bars represent the 95% Wald-type confidence intervals for the predicted average count**

```{r}

## get data right

tick_no2016 <- Tick %>%
  drop_na() %>%
  filter(year != "2016")

# re-order levels to make it easier to interpret effect of fence

tick_no2016$plot <- factor(tick_no2016$plot,
                             levels = c("Partial","Open",  "Total"))


tick_no2016$site <- factor(tick_no2016$site,
                             levels = c("Intermediate", "Arid", "Mesic"))


## best fit model
mod_deoc_no2016 <- glmmTMB(deoc ~ plot + site + plot*year + (1|month),
                           family = nbinom2,
                           data = tick_no2016)



pred <- glmmTMB(deoc ~ plot + site, data = tick_no2016, family = nbinom2)
newdata0 = newdata = unique(tick_no2016[,c("plot","site")])
temp = predict(pred, newdata, se.fit = TRUE, zitype = "response")
newdata$predFE = temp$fit
newdata$predFE.min = temp$fit-1.98*temp$se.fit
newdata$predFE.max = temp$fit+1.98*temp$se.fit

real = ddply(tick_no2016, ~ plot + site, summarize, m = mean(deoc))
newdata$plot <- factor(newdata$plot,
                       levels = c("Total", "Partial", "Open"),
                       labels = c("- Wildlife & Cattle", "Control", "+ Wildlife & Cattle"))
ggplot(newdata, aes(plot, predFE, color = site)) +
  geom_point(size = 3.5) +
  geom_errorbar(aes(ymin = predFE.min, ymax=predFE.max), width = .1, size = 1.5) +
  #geom_point(data=real, aes(x=plot, y=m)) +
  labs(y = "Estimated Average Tick Abundance", x = "", color = "Site") +
  theme_classic() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(face = "bold", size = 14, color = "black"),
        legend.position = c(.83,.8),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold"),
        legend.background = element_rect(color = "black", fill = "white", size = .1),
        legend.title.align = .5) + #,
    scale_color_manual(values =c('tan1','brown1','blue'))
#ggsave("~/Google Drive/Tejon_clean/data_visualizations/Fig16_PredictedAverageCount.jpeg", dpi = 320)

```

