---
title: "Data Analysis v3"
author: "sbsambado"
date: "1/05/2023"
output:
  pdf_document: default
  html_document: default
---

# Manuscript: Variable effects of wildlife and livestock on tick abundance across a climatic gradient
## Authors: Copeland S, S Sambado, D Orr, A Bui, A Swei & H Young
### Component: Data analysis 


**Terminology**

Terminology that may differ between manuscript and data analysis:

*herbivore grazing pressure = plot*
- Open = CW 
- Partial = W, control
- Total = - CW 

*climate aridity level = site*
- Arid
- Intermediate
- Mesic

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload neccessary packages
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(car)
library(knitr)
library(lme4)
library(jtools)
library(effects)
library(psych)
library(gt)
library(MuMIn)
library(kableExtra)
library(MASS)
library(rstatix)
library(here)

## upload data
tick <- read_csv(here("data_analyses", "Tejon_MeanComparisons_Dataset.csv"))
dim(tick) # 1449   11


## add additional information to data

# update sampling effort in these analyses
Tick <- tick %>%
  mutate(sampling_effort = case_when(
    year == "2016" ~ "9",
    year == "2017" ~ "15",
    year == "2018" ~ "15",
    year == "2019" ~ "13"

  ))

Tick <- Tick %>%
  mutate(treatment = case_when(
    year == "2016" ~ "pre",
    year != "2016" ~ "post"
  ))

## format data

Tick$deoc <- as.integer(Tick$deoc)
Tick$site <- as.factor(Tick$site)
Tick$plot <- as.factor(Tick$plot)
Tick$year <- as.factor(Tick$year)
Tick$month <- as.factor(Tick$month)
Tick$sampling_effort <- as.numeric(Tick$sampling_effort) 



```

## RESULTS

### Part 1: summary of study 

**summary statistics**
Sampling from 2016 – 2019 resulted in 2,692 questing ticks. The most common tick species collected via tick drag were Dermacentor occidentalis (n = 2,420, 89.9% of the collected tick population), Ixodes pacificus (n = 206, 7.7%), and Dermacentor variabilis (n = 68, 2.5%) (Supplemental Table 1). Despite extensive effort to collect the juvenile life stages of all tick species, we had limited success (a regionally common problem) and therefore only the adult life stages were included in the analysis. 

The mean of questing tick counts were greatest in 2017 (2.8 +- 5.3), one year post fence placement, followed by 2018 (1.9 +- 3.5) and 2019 (2.1 +- 6.4) with mean counts of  0.22 +- 0.52 ticks in 2016 during exclosure fence construction. When comparing mean tick counts across herbivore grazing pressure treatments throughout the study, CW plots has the lowest (1.05 +- 2.9) followed by W (2.1 +- 4.7) and then – CW (2.6 +- 5.9). Mean tick counts by aridity level were highest in Intermediate (3.37 +- 6.6, n = 1,764), followed by Mesic (1.11 +- 2.04, n = 588), and Arid (0.64 +- 2.72, n = 340).  There was a strong seasonal pattern of mean tick counts with majority of ticks being collected between April through June. The mean of total tick counts aggregated across 2016-2019 were highest in April (4.40 +- 7.66) followed by May (3.51 +- 6.55), and June (2.99 +- 4.55) with the rest of the months (July – March) having a mean total tick count of less than 2.00 (Supplemental Figure 2).

(ALL TICKS)

```{r}

### raw summaries 

Tick %>%
  tally(total)

Tick %>%
   tally(deoc)

Tick %>%
   tally(ipac)

Tick %>%
   tally(deva)

### mean questing ticks

## year
Tick %>%
  group_by(year) %>%
  dplyr::summarise(Mean = mean(total, na.rm = TRUE),
            SD = sd(total, na.rm = TRUE))

## herbivore grazing pressure
Tick %>%
  group_by(plot) %>%
  dplyr::summarise(Mean = mean(total, na.rm = TRUE),
            SD = sd(total, na.rm = TRUE))

## climate aridity
Tick %>%
  group_by(site) %>%
  dplyr::summarise(Mean = mean(total, na.rm = TRUE),
            SD = sd(total, na.rm = TRUE))
 

## month

Tick %>%
  group_by(month) %>%
  dplyr::summarise(Mean = mean(total, na.rm = TRUE),
            SD = sd(total, na.rm = TRUE))
```

### Part 2: effects of herbivore grazing and climate aridity treatments on D. occidentalis counts

(Now only D. occidentalis)


**Effect sizes of herbivore grazing pressure and climate aridity varied in magnitude**
The Kruskal-Wall test showed there was a significant difference between herbivore grazing pressure levels (X^2 = 12698, df = 12, p-value < 0.001) and climatic aridity levels (X^2 = 12698, df = 12, p-value < 0.001) on D. occidentalis counts. For herbivore grazing pressure, the post-hoc Dunn test showed that Open compared with Total (z-test statistic = 5.1, p-value < 0.001) and Partial compared with Open (z-test statistic = -4.4, p-value < 0.001) were significantly different. For climatic aridity levels, the post-hoc Dunn test showed that all three climatic levels were significantly different from each other with Intermediate compared with Arid the most significantly different (z-test statistic = -9.4, p-value < 0.001), followed by Intermediated compared with Mesic (z-test statistic = -5.3, p-value < 0.001), and Arid compared with Mesic (z-test statistic = 4.1, p-value < 0.001). Based on the Kruskall-Wallis test, the effect size for herbivore grazing pressure (H-statistic = 0.0196) was small while the effect size for climatic aridity levels was moderate (H-statistic = 0.060) for D. occidentalis counts.

```{r}
### PLOT (herbivore pressure)
# step 2. run kruskall wallis test
Tick %>% 
  kruskal.test(deoc ~ plot)
# Kruskal-Wallis chi-squared = 12698, df = 12, p-value < 2.2e-16


# step 3. post-hoc test Dunn test
Tick %>%
  dunn_test(deoc ~ plot, p.adjust.method = "bonferroni")## p adj sig for Partial vs Open (2.620560e-05) and Open vs Total (1.281627e-06)

## step 4. whole year plot differences (effect size)
Tick  %>%
  kruskal_effsize(deoc ~ plot)

# effect size: 0.0196908	eta2[H]	small

## based on 
#The results of the Kruskall-Wallis chi-squared test were singicat
#Pairwise comparisons using Dunns test indicate that Partial was significantly different from those of Open (z = -4.4, p adjusted < .001)



### SITE (climatic level)
# step 2. run kruskall wallis test
Tick %>% 
  kruskal.test(deoc ~ site)
# Kruskal-Wallis chi-squared = 12698, df = 12, p-value < 2.2e-16


# step 3. post-hoc test Dunn test
Tick %>%
  dunn_test(deoc ~ site, p.adjust.method = "bonferroni")## p adj sig for all 

## step 4. whole year site differences (effect size)
Tick  %>%
  kruskal_effsize(deoc ~ site)

# effect size: 0.05988807	eta2[H]	small

## based on 
#The results of the Kruskall-Wallis chi-squared test were singicat
#Pairwise comparisons using Dunns test indicate that Partial was significantly different from those of Open (z = -4.4, p adjusted < .001)
```


**Levels between herbivore grazing pressure and climate aridity treatments were significantly different than control set up of partial intermediate**
We found that herbivore grazing pressure, climate aridity context, and year since exclosure fence construction had an impact on D. occidentalis counts with our GLMM models. We will first discuss the impact of individual fixed effects – herbivore grazing pressure or climate aridity – and then evaluate the interactive effects of both stressors on tick counts. For our first research question, does herbivore grazing pressure have an impact on counts, the best fit model included fixed effects herbivore grazing pressure and year with a conditional R2 of 0.26 (fixed effects) and marginal R2 of 0.98 (fixed and random effects). Compared with the model intercept, partial (W) at year 2016 (0 years since exclosure fence construction), only open (CW) was significantly different (estimate = -1.075 ± 0.445, z-value = -2.4, p-value = 0.016) and is expected to have 1.08 lower log counts of D. occidentalis than Partial (W). For the coefficient year, all three years after fence construction (2017, 2018, 2019) were significant with the greatest decrease in D. occidentalis counts occurring in 2018 (estimate = -3.817 ± 0.1625, z-value = -23.48, p-value < 0.001), then 2017 (estimate = -3.422 ± 0.1625, z-value = -21.25, p-value < 0.001), and 2019 (estimate = -1.506 ± 0.16259, z-value = -9.265, p < 0.001). After accounting for the variance explained by fixed effects, the variance explained by random effects in the model was 87% for month and 13% for plot ID.  For our second research question, what are the effects of climatic aridity levels on tick abundance, the best fit model included the fixed effect of an interaction term between climate aridity and year (proxy for how many years exclosure fences have been constructed) with a conditional R2 of 0.31 (fixed effects) and marginal R2 of 0.98 (fixed and random effects). Compared with the model intercept Intermediate at year 2016, both Mesic (estimate = -1.69 ± 0.651, z-value = -2.60, p = 0.009) and Arid (estimate = -1.31 ± 0.549, z-value = -2.38, p = 0.017) were significantly different and had less expect log counts of ticks than Intermediate. Like the results in question one, all years (2017, 2018, 2019) had lower expected log counts of ticks and were significantly different than the year (2016) (see Supplement for additional summaries). The only significant interaction terms between climate aridity level and year were Arid in 2018 (estimate = -.978 ± 0.424, z-value = -2.306, p = 0.02) and Mesic in 2017 (estimate = 1.095 ± 0.537, z-value = 2.038, p = 0.04) with higher expected log counts of ticks in Arid 2018 and lower expected log counts of ticks in Mesic 2017 compared with Intermediate 2016. When we evaluated the effect of the interaction between both fixed effects – herbivore grazing pressure and climate aridity – we found that the interaction between Open (CW) and Arid sites as well as Arid sites were significantly different than Partial (W) Intermediate with a conditional R2 of 0.14 (fixed effects) and marginal R2 of 0.97 (fixed and random effects). The Open (CW) Arid site (estimate = -1.89 ± 0.6788, z-value = -2.78, p-value = 0.005) and Arid site (estimate = -0.964 ± 0.449, z-value = -2.417, p-value = 0.031) alone had lower expected log counts of ticks than Partial (W) Intermediate. 


(DEFAULT INTERCEPTS --> 2016 Partial (W) Intermediate)


####Question 1: What are the effects of herbivore grazing pressure treatments on D. occidentalis tick counts?

```{r}
## Set Intercept : Year 1 Partial

Tick$plot = relevel(Tick$plot, ref = "Partial")


## Model construction -- change fixed effects
mod_1_1 <- glmer(deoc ~ 1 + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)


mod_1_2 <- glmer(deoc ~ plot + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_1_3 <- glmer(deoc ~ plot + year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_1_4 <- glmer(deoc ~ plot*year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick) # failed to converge

## Compare Fixed Effects Models with random held constant

AIC(mod_1_1, mod_1_2, mod_1_3, mod_1_4) # mod_4 lowest, but mod_3 also good, going with mod_3 since mod_4 failed to converge

anova(mod_1_3, mod_1_4) # mod4 better, but going with mod_3 since mod_4 failed to converge

## With best fixed effects, let's change random to see what's best fit model 
## let's see about random effect

mod_1_5 <- glmer(deoc ~ plot + year + offset(sampling_effort) + (1|month),
                 family = "poisson", 
                 data = Tick)


mod_1_6 <- glmer(deoc ~ plot + year + offset(sampling_effort) +  (1|plotID),
                 family = "poisson", 
                 data = Tick) # failed to converge

AIC(mod_1_3, mod_1_5, mod_1_6) # mod3 still better

## Best fit model for this question
summary(mod_1_3)


# marginal = fixed factors, conditional = both fixed and random
r.squaredGLMM(mod_1_3) # marginal = .26, conditional = .98


```

####Question 2: What are the effects of climate aridity treatments on D. occidentalis tick counts?

```{r}
# Set Intercept Year 1 Intermediate
Tick$site = relevel(Tick$site, ref = "Intermediate")

## Model construction -- change fixed effects
mod_2_1 <- glmer(deoc ~ 1 + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)


mod_2_2 <- glmer(deoc ~ site + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_2_3 <- glmer(deoc ~ site + year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_2_4 <- glmer(deoc ~ site*year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)
#model failed to converge with max|grad| = 0.00985153 (tol = 0.002, component 1)

#variance inflation factor - indicates moderate to high multicollinearity of the site, year and site:year 
library(car)
vif(mod_2_4)

#install.packages("brglm2")
library(brglm2)

mod_2_4_2 <- glmer (deoc ~ site*year + offset(sampling_effort) + (1|month) + (1|plotID),
                    family = "poisson",
                    data = Tick,
                    control = glmerControl(optimizer = "bobyqa"))

summary(mod_2_4_2)

#checking optimizers
library(dfoptim)
library(optimx)
allFit(mod_2_4, verbose = F)

## including a BOBYQA optimization allows to model to converge. The result through a low difference in the negative log-likelihoods noted that different optimizers provided similar results in terms of the likelihood of the data 


## Compare Fixed Effects Models with random held constant

AIC(mod_2_1, mod_2_2, mod_2_3, mod_2_4) # mod_4 lowest, but mod_3 also good

anova(mod_2_3, mod_2_4) # mod4 better

## With best fixed effects, let's change random to see what's best fit model 
## let's see about random effect

mod_2_5 <- glmer(deoc ~ site*year + offset(sampling_effort) + (1|month),
                 family = "poisson", 
                 data = Tick) # failed to converge


mod_2_6 <- glmer(deoc ~ site*year + offset(sampling_effort) +  (1|plotID),
                 family = "poisson", 
                 data = Tick)

AIC(mod_2_4, mod_2_5, mod_2_6) # mod4 still better

## Best fit model for this question
summary(mod_2_4)

summary(mod_2_4_2)

summary(mod_2_3)

# marginal = fixed factors, conditional = both fixed and random
r.squaredGLMM(mod_2_4) # marginal = .31, conditional = .98

r.squaredGLMM(mod_2_4_2) # marginal = .31, conditional = .98

r.squaredGLMM(mod_2_3) # marginal = .29, conditions = .98
```

####Question 3: What are the effects of the interaction between herbivore grazing pressure and climate aridity treatments on D. occidentalis tick counts?

```{r}
## Set Intercept Year 1 Intermediate Partial

Tick$site = relevel(Tick$site, ref = "Intermediate")
Tick$plot = relevel(Tick$plot, ref = "Partial")

## Model construction -- change fixed effects
mod_3_1 <- glmer(deoc ~ 1 + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)


mod_3_2 <- glmer(deoc ~ plot + site + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_3_3 <- glmer(deoc ~ plot + site + year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)

mod_3_4 <- glmer(deoc ~ plot*site  + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick)


mod_3_5 <- glmer(deoc ~ plot*year + site*year + offset(sampling_effort) + (1|month) + (1|plotID),
                 family = "poisson", 
                 data = Tick) # failed to converge

## Compare Fixed Effects Models with random held constant

AIC(mod_3_1, mod_3_2, mod_3_3, mod_3_4, mod_3_5) # mod_3 lowest but doesn't answer question like mod 4 does

anova(mod_3_3, mod_3_4) # they actually aren't significantly different so will keep mod_4

## With best fixed effects, let's change random to see what's best fit model 
## let's see about random effect

mod_3_5 <- glmer(deoc ~ plot*site + offset(sampling_effort) + (1|month),
                 family = "poisson", 
                 data = Tick) # failed to conver


mod_3_6 <- glmer(deoc ~ plot*site+ offset(sampling_effort) +  (1|plotID),
                 family = "poisson", 
                 data = Tick)

AIC(mod_3_4, mod_3_5, mod_3_6) # mod4 still better


## Best fit model for this question
summary(mod_3_4) 

# marginal = fixed factors, conditional = both fixed and random
r.squaredGLMM(mod_3_4) # marginal = .14, conditional = .97

```

**Model comparison and diagnostics**

All models

```{r}
### Comparing df, AIC, BIC
model_results <- AIC(mod_1_1, mod_1_2, mod_1_3, mod_1_4,
                     mod_2_1, mod_2_2, mod_2_3, mod_2_4,
                     mod_3_1, mod_3_2, mod_3_3, mod_3_4, mod_3_5)

model_models <- list(mod_1_1, mod_1_2, mod_1_3, mod_1_4,
                     mod_2_1, mod_2_2, mod_2_3, mod_2_4,
                     mod_3_1, mod_3_2, mod_3_3, mod_3_4, mod_3_5)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)
for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$fixed_effects <- c("null", "plot", "plot + year", "plot*year",
                             "null", "site", "site + year", "plot*year",
                             "null", "plot + site", "plot + site + year", "plot*site", "plot*site + site*year")

model_models_full <- kable(model_results, digits = 2, align = "c", 
                      caption = "D. occidentalis counts ~ fixed = fixed_effects column + offset(sampling days),  random effects = (1|month) + (1|plotID)") %>%
  kable_styling() 

model_models_full


```


Best fitting models
```{r}

### Basic model summaries
summary(mod_1_3) # effect of herbivore grazing
summary(mod_2_4) # effect of climate aridity
summary(mod_3_4) # effect of interacting herbivore x aridity


### Comparing df, AIC, BIC
model_results <- AIC(mod_1_3, mod_2_4, mod_3_4)

model_models <- list(mod_1_3, mod_2_4, mod_3_4)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)
for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$variables <- c("herbivore grazing + year", "climate aridity*year", "herbivore*aridity")

model_models_best <- kable(model_results, digits = 2, align = "c", 
                      caption = "D. occidentalis counts ~ fixed = fixed_effects column + offset(sampling days),  random effects = (1|month) + (1|plotID)") %>%
  kable_styling()

model_models_best ## PRINT THIS SUMMARY OUT


### Calculating Akaike weight
# (get relative support for each model by standardizing the likelihood of the model)

## Dredge the model

min(Tick$total, na.rm = TRUE) # 0

# remove NAs
model <- na.omit(Tick)

# build basic poisson model with both herbivore and aridity treatments
big.model <- glm(deoc  ~ plot*site, 
                 family = "poisson",
                 data = model,
                 na.action = na.pass)

# dredge model
sal.dredge <- dredge(big.model)


# format model output
glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  tab_header(title="Model Selection",
             subtitle = "deoc counts ~ plot*site")  %>%
  fmt_number(
    columns = c(1, 6:9),
    decimals = 2
  )

glm.best ## PRINT THIS SUMMARY OUT

# delta = 0 is best model

```
Create output table of model results
```{r}

# export_summs(mod_1_3, mod_2_4, mod_3_4,
#            scale = TRUE,
#            exp = TRUE,
#            error_format = "({statistic}, 95% CI [{conf.low}, {conf.high}], p = {p.value})",
#              to.file = "docx", 
#              file.name = "Tejon_BestFitGLMMs_v1.docx")

```


### Part 3: investigating possible mechanisms beyond the treatment effects

**How do environmental stressors impact population estimates of lizards and mammals, two important vertebrates in the tick system?**
Small mammals and lizards were impacted differently by cattle pressures and climate characteristics with lizard population estimates responded more strongly to climate aridity levels (site) whereas mammals responded more to cattle grazing pressure (plot) (Supplemental Figure 3). Cattle and wildlife presence had a negative relationship with estimated mammal abundance, with mammal abundance highest in the total exclosure plot (-WC). Based on sampling sessions across the entire study, lizard population estimates (n = 1,799) were three times higher than mammal population estimates (n = 561). Despite the smaller sample size (Supplemental Fig), mammal population estimates trend higher in exclosure plots Total (n = 196) and Partial (n = 305) compared to Open (n = 61) plots but were not statistically significant. Whereas lizard population estimates were significantly different by site type (Wilcoxon rank sum: W = 11, p-value = 0.0131) and non-significantly different by plot type. From the Wilcoxon test, the median of lizard population estimates by site types were doubled in Intermediate (n = 134) compared to Arid (n = 70). Mean? Shannon diversity indices of mammal populations increased with exclosure intensity (Open = 0.607, Partial = 0.892, Total = 0.927) but was not statistically significant.

Step 1. Import & tidy dataset

```{r import and tidy}

pop <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mamliz_popestimates.csv")

# rename columns to be consistant
names(pop)[5] <- "mam_estimate"

# change data structure for analyses
pop$site <- as.factor(pop$site)
pop$plot <- as.factor(pop$plot)

```

Step 2. Compare means of population estimates

2A. Lizards
```{r lizards}
# lizards were only sampled at two sites, so will use a two sample t-test if normal

# check normality
qqPlot(pop$liz_estimate) # looks normal, but..
shapiro.test(pop$liz_estimate) # p < 0.05, so data are not normal

# check variance
leveneTest(pop$liz_estimate ~ pop$site) #  p < 0.05, F = 7.4; variances are different

# non-parametric Mann-Whitney U is testing the shape of the two distributions are different

# wilcox rank sum test with continuity correction (non-equal variances)
wilcox.test(liz_estimate ~ site, data = pop) #W = 11, p-value  = 0.01306

# p < 0.05, reject H0 that means are equal to 0

## lizard population estimates are different at Intermediate and Arid sites

```

2B. Mammals
```{r mammals}

# mammal populations did not significantly differ across sites or plots but that may be due to low sample numbers with kruskal wallis

# let's do open vs total plots --> mann-whitney U test

pop_sub <- subset(pop, pop$plot != "Partial")
wilcox.test(mam_estimate ~ plot, data = pop_sub)
# W = 11, p-value = 0.05488 (trending significant)
```


**Are vertebrate population estimates correlated with tick counts?**

We found that lizard population estimates were significantly and positively correlated with tick counts (p-value < 0.001, rho = 0.17) whereas mammal population estimates were not significant or correlated (p-value = 0.37, rho = -0.04). 

Step 1. Import & tidy dataset

```{r import and tidy}
FINAL <- read_csv("~/Google Drive/Tejon_clean/data_analyses/Tejon_MixedModels_Dataset.csv")

# drop N/A values
FINAL <- FINAL[which(FINAL$mam_estimates != "#N/A"),]
FINAL <- FINAL[which(FINAL$liz_estimates != "#N/A"),]

# check it worked
unique(FINAL$mam_estimates)
unique(FINAL$liz_estimates)

str(FINAL)
```

Step 2. Spearman's rank correlation

```{r spearman rank correlation}
# select subset of data to use for test & visualize
FINAL_cor <- subset(FINAL, select = c(6,20:21))

# change structure of data for analyses
FINAL_cor$total <- as.numeric(FINAL_cor$total)
FINAL_cor$mam_estimates <- as.numeric(FINAL_cor$mam_estimates)
FINAL_cor$liz_estimates <- as.numeric(FINAL_cor$liz_estimates)

# visualize data
pairs.panels(FINAL_cor)

## run correlation test

# lizards
cor.test(FINAL_cor$total,FINAL_cor$liz_estimates,
         method = "spearman")
# p-value = 1.755e-05
# rho = 0.1687151

# mammals
cor.test(FINAL_cor$total,FINAL_cor$mam_estimates,
         method = "spearman")
# p-value = 0.3669
# rho = -0.03569831 

```


